---
title: "A paper in linguistics"
author: "A researcher"
date: today
lang: en
format: typst
df-print: kable
bibliography: references.bib
csl: apa
---

```{=typst}
    #import "@preview/phonokit:0.5.9": *
    #import "@preview/synkit:0.0.3": *
```

```{r}
#| message: false
#| warning: false
#| echo: false

library(tidyverse)
library(Fonology)

d <- read_csv("../data/danish-1.csv")
```

# Data analysis

We have already loaded a data file above (invisible here), so we can now easily run R code internally. For example, the mean reaction time in our data is `r mean(d$RT) |> round(2)`ms. For `lig` specifically, it's `r d |> filter(Affix == "lig") |> summarize(rt = mean(RT)) |> pull(rt) |> round(2)`ms.

We can print a proper table by adding `df-print: kable` to our preamble. This will turn any data frame/tibble outputs into a _real_ table.

:::{#tbl-table1}
```{r}
#| echo: false
d |> 
  summarize(`Mean RT (ms)` = mean(RT) |> round(2), .by = Affix)
```

A nice-looking table from our data
:::

## Figure

As usual, we can also create a nice figure directly within our `Qmd` file.


```{r}
#| message: false
#| warning: false
#| echo: false
#| fig-cap: "A figure in our document"
#| fig-width: 6
#| fig-height: 4
#| fig-dpi: 500
#| out-width: 60%
#| label: fig-myfig

ggplot(
  data = d |>
    filter(
      Participant %in% c("Part_5", "Part_15", "Part_8"),
      Affix %in% c("bar", "ede")
    ),
  aes(
    x = Affix |> fct_reorder(RT, .fun = mean),
    y = RT,
    color = Participant
  )
) +
  stat_summary(aes(group = Participant),
    geom = "line",
    linetype = "dashed",
    alpha = 0.5,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_cl_normal,
    geom = "errorbar", width = 0,
    position = position_dodge(width = 0.7),
    linewidth = 3,
    # color = "slateblue",
    alpha = 0.2
  ) +
  stat_summary(position = position_dodge(width = 0.7)) +
  geom_label(
    data = d |>
      filter(
        Participant %in% c("Part_5", "Part_15", "Part_8"),
        Affix %in% c("bar")
      ) |>
      summarize(
        RT = mean(RT),
        .by = c(Affix, Participant)
      ),
    aes(label = Participant),
    position = position_dodge(width = 0.7),
    # NOTE: Or you can use position_nudge(x = ...) if
    # you prefer to have the labels *next* to the means
    # family = "Futura"
  ) +
  theme_classic(
    base_size = 13,
    base_family = "Futura"
  ) +
  theme(
    legend.position = "none"
  ) +
  labs(
    x = "Affix",
    y = "Reaction time (ms)",
    title = "Danish data"
  ) +
  # NOTE: If you want to change colours:
  scale_color_manual(
    values = c("black", "darkorange2", "slateblue")
  )


```



# Linguistics

Phonetic transcription can be done using the `Fonology` package [@fonology]. The package supports English, French, Italian, Portuguese and Spanish. For example, the word _comfortably_ is [`r ipa("comfortably", lg = "en")`]. Check `gdgarcia.ca/fonology` for more info on the package.

With Typst, we can load `phonokit` [@phonokit_garcia] and `synkit` [@synkit_garcia], which in turn allow us to produce virtually any linguistic representation, like the prosodic word in @fig-word1 and the tree in @fig-tree1.


:::{#fig-word1}
`#word("('k \\ae.n@).d@", scale: 0.8)`{=typst} 

Prosodic structure in Quarto
:::

Check the `Qmd` file for the code, but it's pretty minimal. For example, @fig-word1 is generated with `#word("('k \\ae.n@).d@")`, which uses the familiar `tipa` reference for phonetic symbols (but you can technically add symbols directly into the document, since Unicode is fully supported in Typst).


:::{#fig-tree1}
`#tree("[TP [NP the students_i] [T' [T pres] [VP [NP t_i] [V' [V like] [NP cats]]]]]")`{=typst}

A syntax tree
:::

Likewise, in-line arrows and numbered examples are available. 

:::{#fig-inline}
```{=typst}
  #move(
      "[CP Who do you think [(CP)[TP<who>ate the chocolate?]]]",
      arrows: (
        (from: "who2", to: "who1"),
      ),
      protect: true,
    )
```
In-line movement with the function `#move()` in `synkit`
:::

Here's another example, this time printing vowels and consonants in English.

:::{#fig-vowels}
`#vowels("english", scale: 0.5)`{=typst}

Vowel inventory in English
:::

The `#consonants()` function is also handy: it comes with some preset languages, but you can also set up your own inventory.

:::{#fig-consonants}
`#consonants("english", affricates: true, simplify: true, scale: 0.6)`{=typst}

Consonant inventory in English
:::


